### 2017-01-21 main.tex

\documentclass[11.5pt,two column]{llncs}

\setcounter{tocdepth}{3}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb, amsmath}
\usepackage{array,url}
\usepackage{authblk}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tikz}\usetikzlibrary{arrows}
\usepackage{times}
\usepackage{url}
\usepackage{xcolor}

% change the appearance of the tikz arrow for the argumentation networks
% and some other settings to make all the graphs look similar
\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}

\addtolength{\oddsidemargin}{-0.85in}
\addtolength{\evensidemargin}{-0.85in}
\addtolength{\textwidth}{1.7in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.3in}
\setlength{\columnsep}{0.3in}
\newcommand{\Mark}[1]{\textsuperscript{#1}}

% TODO command
\newcommand\todo[4][]{%
	\ifthenelse{\equal{#1}{resolved}}{%
		% nothing
	}{%
		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
	}%
}

% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
% \renewcommand\todo[4][]{}

% use like this:
% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex

\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}

\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}\selectfont}

\begin{document}
\twocolumn[{%
 \centering
 \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
 \large Marc van Zee\Mark{1},
        Floris Bex\Mark{2},
        and Sepideh Ghanavati\Mark{3}\\[1em]
 \normalsize
 \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
 \end{tabular}\\[3em] % some more space after the title part
}]

\begin{abstract}
Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks. The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders. While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model. For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured. In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling. We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL). We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
\end{abstract}

\newpage

\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering

\input{introduction}
\input{background}
\input{preview}
\input{gmas}
\input{formalframework}
\input{tool}
\input{methodology}
\input{relatedwork}
\input{conclusion}

\section*{Acknowledgments}
Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.

\bibliographystyle{abbrv}
\bibliography{references}

\newpage

\onecolumn
\appendix

\input{transcripts}

\end{document}