### 2017-01-21 tool.tex

\section{The RationalGRL Tool}
\label{sect:implementation}