### 2017-01-23 main.tex
minor typo - test

\documentclass[11.5pt,two column]{llncs}

\setcounter{tocdepth}{3}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb, amsmath}
\usepackage{array,url}
\usepackage{authblk}
\usepackage{caption}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{subcaption}\captionsetup{compatibility=false}
\usepackage{tabularx}
\usepackage{tikz}\usetikzlibrary{arrows, backgrounds}
\usepackage{times}
\usepackage{url}
\usepackage{xcolor}

% change the appearance of the tikz arrow for the argumentation networks
% and some other settings to make all the graphs look similar
\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}

% Tikz stuff for arguments with GRL
\tikzset{
  argNodeIN/.style={shape=rectangle,draw,fill=white,align=center},
  argNodeOUT/.style={shape=rectangle, draw=black!60, text=black!40,fill=white},
  traceNodeRED/.style={circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt},
  traceNodeGREEN/.style={circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt},
  attackLink/.style={draw=black,->,line width=1mm},
  traceLinkRED/.style={draw, densely dotted,-,line width=0.5mm, draw=red},
  traceLinkGREEN/.style={draw, densely dotted,-,line width=0.5mm, draw=green},
  CQLink/.style={line width=0.5mm, dotted,->},
  ldots/.style={draw=none,rotate=90,font=\fontsize{20}{22.4}\selectfont},
  grl/.style={draw=none},
  disabled/.style={opacity=0.5}
}
\newcommand{\argtext}[2]{
\begin{tabularx}{0.2\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}

\addtolength{\oddsidemargin}{-0.85in}
\addtolength{\evensidemargin}{-0.85in}
\addtolength{\textwidth}{1.7in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.3in}
\setlength{\columnsep}{0.3in}
\newcommand{\Mark}[1]{\textsuperscript{#1}}

% TODO command
\newcommand\todo[4][]{%
	\ifthenelse{\equal{#1}{resolved}}{%
		% nothing
	}{%
		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
	}%
}

% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
% \renewcommand\todo[4][]{}

% use like this:
% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This will show only in the tex

\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}

\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}\selectfont}

\begin{document}
\twocolumn[{%
 \centering
 \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
 \large Marc van Zee\Mark{1},
        Floris Bex\Mark{2},
        and Sepideh Ghanavati\Mark{3}\\[1em]
 \normalsize
 \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
 \end{tabular}\\[3em] % some more space after the title part
}]

\input{abstract}

\newpage

\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering

\input{introduction}
\input{background}
\input{preview}
\input{gmas}
\input{formalframework}
\input{tool}
\input{methodology}
\input{relatedwork}
\input{conclusion}

\section*{Acknowledgments}
Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.

\bibliographystyle{abbrv}
\bibliography{references}

\newpage

\onecolumn
\appendix

\input{transcripts}

\end{document}