### 2017-10-29 main.tex

\documentclass[twocolumn]{svjour3}

\setcounter{tocdepth}{3}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[fleqn]{amsmath} % flush equations left
\usepackage{amssymb}
\usepackage{array,url}
\usepackage{authblk}
\usepackage{caption}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{multicol}
\usepackage{multirow}
%\usepackage{sectsty}
\usepackage{subcaption}\captionsetup{compatibility=false}
\usepackage{tabularx}
\usepackage{tikz}\usetikzlibrary{arrows, backgrounds}
\usepackage{times}
\usepackage{url}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{changepage}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\newcommand{\var}[1]{{\ttfamily#1}}% variable

% change the appearance of the tikz arrow for the argumentation networks
% and some other settings to make all the graphs look similar
\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}

% Tikz stuff for arguments with GRL
\tikzset{
  argNodeIN/.style={shape=rectangle,draw,fill=white,align=center},
  argNodeOUT/.style={shape=rectangle, draw=black!60, text=black!40,fill=white},
  traceNodeRED/.style={circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt},
  traceNodeGREEN/.style={circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt},
  attackLink/.style={draw=black,->,line width=0.5mm},
  traceLinkRED/.style={draw, densely dotted,-,line width=0.5mm, draw=red},
  traceLinkGREEN/.style={draw, densely dotted,-,line width=0.5mm, draw=green},
  CQLink/.style={line width=0.5mm, dotted,->},
  ldots/.style={draw=none,rotate=90,font=\fontsize{20}{22.4}\selectfont},
  grl/.style={draw=none},
  disabled/.style={opacity=0.5}
}
\newcommand{\argtext}[2]{
\begin{tabularx}{0.2\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}
\newcommand{\smallargtext}[2]{
\begin{tabularx}{0.25\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}
\newcommand{\longargtext}[2]{
\begin{tabularx}{0.5\textwidth}{X}
\ifx&#1&%
\else
\textbf{(#1)}
\fi
#2
\end{tabularx}
}

% rationale
\newcommand{\rationale}{\emph{Rationale and example:}}

%\addtolength{\oddsidemargin}{-0.85in}
%\addtolength{\evensidemargin}{-0.85in}
%\addtolength{\textwidth}{1.7in}
%\addtolength{\topmargin}{-.7in}
%\addtolength{\textheight}{1.3in}
%\setlength{\columnsep}{0.3in}
\newcommand{\Mark}[1]{\textsuperscript{#1}}

% TODO command
\newcommand\todo[4][]{%
	\ifthenelse{\equal{#1}{resolved}}{%
		% nothing
	}{%
		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
	}%
}

\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

%\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
%\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}

%\sectionfont{\fontsize{12}{15}\selectfont}
%\subsectionfont{\fontsize{10}{15}\selectfont}

\begin{document}\sloppy

\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
\author{Marc van Zee \and Floris J. Bex \and Sepideh Ghanavati}
\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl} \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu}}

\maketitle

\input{abstract}

\input{introduction}
\input{background}
\input{gmas}
\input{overview}
%\input{examples}
\input{formalframework}
\input{methodology}
%\input{tool}
\input{discussion}

\bibliographystyle{abbrv}
\bibliography{references}

\newpage

\onecolumn
\appendix

\input{designprompt}
\newpage
\input{transcripts}
\newpage
\input{tool-screenshots}

\end{document}