### 2018-02-16 Response to reviewers.txt
Response to reviewers & small changes

%*Editor: Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.

%FB: perhaps the editor would like us to add more references to RE and REJ? It is true that the majority of references is now outside of RE. **Marc, Sepideh**

-----------------------
%R1: The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema. Critical questions are something that are used to identify arguments and attack links, but  I don't see them as a concept to be included in the meta model. They are part of the methodology used to build the goal model rather concepts of the language.

%FB: **Marc**

-----------------------
%R1: The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool. 
%R2: I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?

Reviewer 1 argues that the algorithms are fairly simple. We agree that many algorithms are very much the same, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two. Reviewer 2 asks whether the automated reasoning can be handled by propositional logic instead of the algorithms. The answer is that propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE. Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic. The idea is that the formalization is needed for the algorithms, but that just providing a formal semantics of GRL does not directly tell you how to compute/reason with GRL models.

%FB: **Marc, can you work on shortening sections 5.4.1-5.4.3?**

-----------------------
%R1: The methodology and the tools looks quite preliminary and without a real validation of the approach. I think this is the weakest part of the paper. Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word. The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.
%R2: I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable. However, this remains to be shown yet in an industrial environment. Connecting them with an existing approach for goal modeling is also a pro in this respect.

%FB: Depending on the editor's answer we can leave this as is in the paper. However, we do need to come up withy a satisfactory example in the response to reviewers. Also, I think we need to include some references on empirical work that has been done on goal modelling ***Sepideh?***

-----------------------
%R2: the statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.

Reviewer 2 comments that the connection to an existing approach, GRL is commendable, but that we should mention in the abstract and the introduction that we use a simplified version of GRL. We have included in the introduction that we use a simplified version. Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution. This simplification does not in any way influence the case study or any of the other results. Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.

%R2:I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL.
In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

%FB: **Floris to write on GRL beliefs in 7.1**

-----------------------
%R2: How would you compare this specific approach with the old gIBIS?
%FB: **Floris to write on gIBIS in 7.1**

-----------------------
%R2: Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?
%FB: **Floris to write on this in 7.1**

-----------------------
%R2: I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.
%FB: **Floris to write on this in 5.2**

-----------------------
%R2: Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled. Validation would mean that a goal model makes sense in the real world as having the "right" goals and relationships between them.
"Verification" in 6.2 appears to stand for evaluation. This cannot be "verified".

%FB: **Marc**

-----------------------
%R2: I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".

%FB: **Sepideh, Marc**

-----------------------
%R2: please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE.

%FB: who originally wrote these sentences? **Marc, Sepideh**

-----------------------
%R2: Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.

%FB: Reference to figure is line 56 p. 19 **Marc**
%FB: A quick scan in my TeX environment with UK spelling shows me no spelling mistakes, and I believe gramatically there's not too much wrong with the paper either, but let's check **Marc, Floris, Sepideh**