### 2018-05-04 Response to reviewers.txt
final comments FB fixed

######## EDITOR ########

===== COMMENT #1
Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.

===== REPLY #1
Currently 6 of the 44 references in the article come from either REJ or the RE conference [11,24,26,32,43,47]. We have searched through the journal and conference database for more references, and we found 4 more articles that are related to our work [20,22,27,39]. We have added a discussion of these articles to the related work section, at the beginning of the paragraph "Requirements Engineering". This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.

######## REVIEWER #1 ########

We thank the reviewer for the insightful comments. Especially the comment about the missing validation was useful since it motivated us to perform an small-scale in-house user study, which allowed us to strengthen our claims for the usefulness and usability of our framework. We’ve replied to each separate comment below.

===== COMMENT #1
The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema. Critical questions are something that are used to identify arguments and attack links, but  I don't see them as a concept to be included in the meta model. They are part of the methodology used to build the goal model rather concepts of the language.

===== REPLY #1
We thank the reviewer for pointing this out. We agree critical questions should not be part of the metamodel, and we have removed them accordingly. We also have updated the text accordingly.

===== COMMENT #2
The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool.

===== REPLY #2
We agree that many algorithms are similar, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.

===== COMMENT #3
The methodology and the tools looks quite preliminary and without a real validation of the approach. I think this is the weakest part of the paper. Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word. The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.

===== REPLY #3
We agree with the reviewer that a real validation of the approach is a necessary next step. However, we would like to point the reviewer to the case study in section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com). We do agree that further validation of the tool and method is necessary, preferably in an industrial setting. However, such an industrial strength case study is a significant amount of work (see for example Estrada et al. (2006 "An empirical evaluation of the i* framework in a model-based software generation environment". They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.

Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content. Still, we agree that having no validation at all is limiting, so we have performed a small-scale user study in order to provide an initial validation of our framework. This empirical study was performed amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.

We have added a section "Validation" (Section 7) where we describe the results of the validation.

######## REVIEWER #2 ########

We thank the reviewer for reading our article in detail, and providing many insightful and useful comments. We’ve replied to each separate comment below.

===== COMMENT #1
I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable. However, this remains to be shown yet in an industrial environment.

===== REPLY #1
We have added a section "Validation" (Section 7) where we describe the results of a validation. Please see REPLY #3 to REVIEWER #1 for more details.

===== COMMENT #2
The statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.

===== REPLY #2
We have included in the abstract and introduction that we use a simplified version. Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution. This simplification does not in any way influence the case study or any of the other results. Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would not change the essence of our formalism.

===== COMMENT #3
I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL. In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

===== REPLY #3
We have added a comparison between arguments in RationalGRL and beliefs in GRL in the related work section, at the beginning of the paragraph "Goal Modelling".

The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below). However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.

===== COMMENT #4 & #5
How would you compare this specific approach with the old gIBIS?
Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?

===== REPLY #4 & #5
We have added some discussion in the 2nd paragraph of section 8.1-Design Rationale. A brief comparison with, for example gIBIS, is already in section 8.1-Design Rationale (reference [38] includes gIBIS). We felt that an extensive comparison between the ideas behind goal modelling, which we extend with RationalGRL, and design rationale diagramming methods like gIBIS is not within the extent of this paper, where the focus is more on the connection between "normal" goal modelling (GRL) and extended "argumentative" goal modelling (RationalGRL).

===== COMMENT #6
I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?

===== REPLY #6
Propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE. Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic. The idea is that the formalization is needed for the algorithms, but that just providing a formal semantics of GRL does not directly tell you how to compute/reason with GRL models.

===== COMMENT #7
I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.

===== REPLY #7
While there is extensive work on less "binary" frameworks for argumentation (see e.g below references 2 and 3), we feel that a discussion on this would not be suitable for the current paper (as it is more of an AI subject than relevant for RE). For an extensive discussion on the relation between the formal argumentation proposed in our article and argumentation in reality, we refer the reviewer to reference [4] below. Note that we have opted not to include these references in the paper, as we are already at 50 references and want to focus more on an RE/SE audience. If the reviewer desires, however, we can add the references to the article.

===== COMMENT #8
Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled. Validation would mean that a goal model makes sense in the real world as having the "right" goals and relationships between them.

===== REPLY #8
We agree that the term “valid” is confusing in this respect, since we do not want to express that the goal model makes sense in the real world. Instead, we’d like to express that certain criteria are fulfilled. Therefore, we’ve renamed “valid” to “consistent”.

**TODO(marcvanzee): Implement this renaming in the article.

===== COMMENT #9
"Verification" in 6.2 appears to stand for evaluation. This cannot be "verified".

===== REPLY #9
We have renamed “verification” to “demonstration” and slightly reworded the sentence.

===== COMMENT #10
I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".

===== REPLY #10
We agree that the terminology is wrong. We have renamed the methodology to "The RationalGRL development process", since it defines the steps on how to create a RationalGRL model. We have renamed this throughout the entire article.

===== COMMENT #11
please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE

===== REPLY #11
We agree with the reviewer that the two mentioned sentences are confusing. We have improved this as follows:
- We have rewritten the first sentence of the abstract as follows: "Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks."
- We have deleted the first sentence of the introduction, since we feel that it is not necessary to explain the definition of Requirements Engineering in the Requirements Engineering Journal.

===== COMMENT #12
Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.

===== REPLY #11
We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.

######## REFERENCES ########

[1] Estrada, H., Rebollar, A. M., Pastor, O., & Mylopoulos, J. (2006, June). An empirical evaluation of the i* framework in a model-based software generation environment. In International Conference on Advanced Information Systems Engineering (pp. 513-527). Springer, Berlin, Heidelberg.
[2] A. Hunter and M. Thimm (2017) Probabilistic Reasoning with Abstract Argumentation Frameworks, Journal of Artificial Intelligence Research, 59:565--611. 
[3] Grossi, D., & Modgil, S. (2015, July). On the Graded Acceptability of Arguments. In IJCAI (pp. 868-874).
[4] van Eemeren, F.H., & Verheij, B. (2017). Argumentation Theory in Formal and Computational Perspective. IFCoLog Journal of Logics and Their Applications 4 (8), 2099-2181.

[5] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari. "Belief revision and argumentation theory." Argumentation in artificial intelligence. Springer, Boston, MA, 2009. 341-360.
[6] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
[7] Paglieri, Fabio, and Cristiano Castelfranchi. "Revising beliefs through arguments: Bridging the gap between argumentation and belief revision in MAS." International Workshop on Argumentation in Multi-Agent Systems. Springer, Berlin, Heidelberg, 2004.