### 2018-05-06 survey.tex
survey

$\ $\\
\newpage
\section{Modeling Task - Traffic Controller}
\label{sect:survey}

A group of three engineers (E1, E2, and E3) is tasked with designing a traffic flow simulation program. They receive a task specification, focusing on how the simulator will handle the generation of cars in the simulation. Below is an excerpt discussing the early requirements of the system. You are asked to model this discussion in as much detail as possible in the RationalGRL tool (http://www.rationalgrl.com), using all the tools RationalGRL puts at your disposal (GRL elements, critical questions, arguments).

As the RationalGRL prototype does not yet have a save functionality, please save a screenshot of your final model.\\

\begin{tabular}{|p{10mm}|p{140mm}|}
\hline
\textbf{Person} & \textbf{Transcript}\\
\hline
E1 & Let's start by adding the actor "Traffic Simulator". Do you think we need to add more actors?\\
\hline
E2 & I don't think so. Let's start with the softgoals and derive the main goals and tasks from that.\\
\hline
E3 & The task description states that the simulator should be easy to use.\\
\hline
E2 & I'm not sure that we should include that. It seems to me this is redundant since any UI should be easy to use.\\
\hline
E1 & No I disagree, for some UI's it is more important than others, and the fact that they mention it explicitly in the task description means we should probably add it anyway.\\
\hline
E3 &I agree, let's keep it in, it seems necessary.\\
\hline
E1 & Ok so we have one softgoal 'easy to use'. Anything else?\\
\hline
E2 & Here it say the simulation should be realistic, and the design should be simple. I guess those are two softgoals as well.\\
\hline
E3 & Ok sounds good. So one goal of the simulation is to Generate cars. Shall we just add that as a goal?\\
\hline
E1 & Yes let's do that.\\
\hline
E2 & I think we then have two main tasks for that goal: The simulator should be able to create new cars in the simulation, and it should be able to wrap the cars when they go out of the screen.\\
\hline
E3 & So Generate cars AND-decomposes into those two tasks?\\
\hline
E1 & I don't think so actually. I think it is a XOR-decomposition. The simulation either creates new cars or wrap them when they are being reused.\\
\hline
E2 & Hmm makes sense. But then I think we should rename Wrap cars to Keep same cars.\\
\hline
E3 & Yes that is better, then Create new cars and Keep same cars are the two alternatives for the goal Generate cars.\\
\hline
E1 & Hmm what do you think about the contribution links? Should we add any?\\
\hline
E2 & Well I think if we create new cars, the simulation will be more realistic because it becomes less static. At the same time it will probably make the design less simple.\\
\hline
E3 & Yes so create new cars contributes positively to the realistic design, and negatively to simple design?\\
\hline
E2 & Hmm now I think about it, actually I think think creating new cars doesn't have much effect on the realism of the simulation.I don't think the user will notice the difference.\\
\hline
E1 & Why not? If you create new cars it changes the flow of the traffic, so I think the user will definitely notice this.\\
\hline
E2 & Okay well, ok maybe. Yes I think you are right. Let's keep the contribution link in then.\\
\hline
\end{tabular}