### 2018-09-17 Response to reviewers2.txt
fixed a typo

Thank you for your considerations.

Reviewer 1 proposes to accept the paper. Reviewer 2 argues that the new section 7 should be revised - we agree with the second reviewer that the style and wording of the section was too informal and extended and sharpened the discussion (see also reply 2 below). Furthermore, reviewer 2 argues that they would like to see more concrete revisions based on some of the points raised in their first review. We have extended the paper, and below we discuss these points from the original review in more detail.

<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< REPLIES FOR REVIEW #2 >>>>>><<<<>>>>>>>>>>>>>>>>>>>>>>
<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

===== COMMENT #1
In the course of the revision, several of my suggestions for improvement have simply been refused. Hence, I do not consider the improvements sufficient.

===== REPLY #1
We have reconsidered all of your suggestions for improvements in the first revision, and did another cycle of improvements of our paper based on your suggestions. We have replied to those suggestions below.

===== COMMENT #2
The newly added Section 7 is unacceptable from a scientific point of view. First, it is called a "small user evaluation". Then it has a subsection entitled "Experimental setup", but there are no hypotheses. Instead, there is another subsection entitled "Analysis", which starts with "Although our empirical evaluation ...". And why is there another subsection about usability of the tool? If that was the concern, a usability test or a heuristic evaluation would have had to be done.

===== REPLY #2
We have significantly reworked section 7, going through the usual steps for an empirical evaluation and including extra statistical results. The section now includes the hypotheses that have been tested, results, a brief statistical analysis of the results, possible threats to validity and a discussion.

<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
<<<<<<<<<<<<<<<<<<<<<<< REVISION OF REPLIES FOR REBUTTAL #1 >>>>>>>>>>>>>>>>>>>>>>>>>>>>
<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

===== ORIGINAL COMMENT #1
I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable. However, this remains to be shown yet in an industrial environment.

===== REPLY #1
Here, we reiterate our argument from the original response to reviewers.

"We agree with the reviewer that a real validation of the approach is a necessary next step. However, we would like to point the reviewer to the case study in Section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com). We do agree that further validation of the tool and method is necessary, preferably in an industrial setting. However, such an industrial strength case study is a significant amount of work (see for example Estrada et al. (2006 "An empirical evaluation of the i* framework in a model-based software generation environment". They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.

Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content. Still, we agree that having no validation at all is limiting, so we have performed a small-scale empirical study in order to provide an initial validation of our framework. This empirical study was performed amongst 16 participants, each with a computer science background."

===== ORIGINAL COMMENT #3
I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL. In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

====== REPLY #3
In our first revision, we already added a brief comparison between arguments in RationalGRL and beliefs in GRL in the related work section (under "Beliefs in GRL"). In this revision, we have done two things. First, we have moved the discussion on beliefs in GRL to "related work" under "goal modeling" and further extended and sharpened our discussion and comparison with GRL beliefs. Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e. given indicated conflicts between beliefs compute ways in which beliefs can be revised).

Second, we have added a new discussion under “Belief Revision“ in section 8.2 (Future work),  where we discuss belief revision and the possibilities this may offer for future research. A more detailed comparison between RationalGRL and a framework based on belief revision would require us to develop a new belief revision framework connected to GRL and then model, for example, the case study or an extended example using this new belief revision framework for GRL. This would be interesting material for a new paper

===== COMMENT #4 & #5
How would you compare this specific approach with the old gIBIS?
Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?

===== REPLY #4 & #5
In the first revision, we added some discussion in the 2nd paragraph of section 8.1-Design Rationale, noting that a comparison with gIBIS is contained in this discussion (reference [38] includes gIBIS). In this revision, we have made this connection explicit. We have clearly indicated that we combine ideas from DR and goal modeling.

Furthermore, in this revision we have given new reasons (under 8.1, second par, "There are two reasons for this....") for why we connect to goal diagramming methods, even though the trend in DR is to not specifically diagram reasons.

===== ORIGINAL COMMENT #6
I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?

===== REPLY #6
In the original reply, we indicated that propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE. Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic. The idea is that the formalization is needed for the algorithms, but that just providing a formal semantics of GRL does not directly tell you how to compute/reason with GRL models.

In this revision, we have added a paragraph to the introduction of section 5, where we explain that we are in fact doing automatic reasoning by implementing the argumentation semantics, but that, from a practical point of view, we have chosen to provide algorithms instead of a propositional logic deduction system.

===== COMMENT #7
I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.

===== REPLY #7
In our original reply, we gave references to less "binary" frameworks for argumentation but did not include them in the revised paper. In this revision, we have included a brief discussion on these frameworks (last paragraph of 5.2), and we also refer to work that discusses the relation between "binary" argumentation and argumentation in the real world in more detail (Rahwan et al. 2010, also see the added footnote X).

===== ALL OTHER ORIGINAL COMMENTS WERE SMALLER COMMENTS (STYLE, TYPO, WORDING) WHICH WE ADDRESSED IN THE ORIGINAL REVISION